---
title: "Atlantic herring (\\protect\\textit{Clupea harengus}) \\linebreak Snapshot Ecosystem & Socioeconomic Profile "
format: 
  pdf:
    margin-left: "0.5in"
    margin-right: "0.5in"
    margin-top: "0in"
    margin-bottom: "0.5in"
    fontsize: "10pt"
    linestretch: 1
    indent: true
    include-in-header: template.tex    
    # keep-tex: true
    papersize: letter
latex_engine: xelatex
params:
  widths: !expr c(0.9, 0.75, 2.5, 3.5)
  img1: !expr here::here("05_images/herrmap.png")
  img2: !expr here::here("05_images/new_condition_resize2.png")
  img3: !expr here::here("05_images/noaa_logo.png")
  tbl_file: !expr here::here("06_docs/table_herring.csv")
  img_dir: !expr here::here("05_images")
  terminal_year: 2024
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#here::i_am("inst/herringESP/06_docs/snapshot_template.qmd")
#devtools::load_all()
```

\vspace{-2cm}

\section{Summer `r params$terminal_year`}

This is a short-form update to the full Ecosystem and Socioeconomic Profile \[1\] highlighting the recent status of environmental and ecological factors. Atlantic herring is an important and valuable New England stock fished primarily by commercial vessels for use as bait (for lobster). The stock is currently overfished but not subject to overfishing.

\vspace{-0.25cm}

::: {layout="[ [0.3, 0.025, 0.675], [1], [0.575, 0.025, 0.4]] "}

:::{}
![](`r params$img1`)
\footnotesize\raggedright Figure source: [https://www.fisheries.noaa.gov/new-england-mid-atlantic/ecosystems/fisheries-habitat-northeast-us-shelf-ecosystem](https://www.fisheries.noaa.gov/new-england-mid-atlantic/ecosystems/fisheries-habitat-northeast-us-shelf-ecosystem)
\normalsize
:::

:::{}
\hfill
:::

::: {}
\section{Recent highlights}

\subsection{2025 Research Track Stock Assessment}

-   Explored a recruitment index from seabird diet data \[2\]
-   Developed indicators of predation by haddock \[3\], food availability \[4\], and temperatures experienced by larvae \[5\] to test as ecosystem covariates for recruitment but none significantly improved the model \[6\]

\subsection{Fishing community observations [7]}

-   Market processes: increased reliance on menhaden due to declining and inconsistent herring catch, reduced quotas, higher fuel prices, river herring bycatch
-   Ecological concerns: warming, changing zooplankton and forage base, haddock predation, altered predator-prey interactions

\subsection{Commercial Fishery}

-   Reduced participation, particularly of larger vessels
-   Broader market impacts include switch to alternative sources like frozen herring and menhaden

\subsection{Management}

-   Still in a period of substantially reduced catch limits
-   Frequently changing ABC and sub-ACLs across the 4 management areas
-   Several extensions and revisions to the target rebuilding date, currently 2031

\subsection{Ecosystem}

-   Age 3+ adults migrate to the Gulf of Maine for summer/fall spawning.
-   Haddock predation on eggs is decreasing
-   Development depends on appropriately sized zooplankton prey at the right time in lifecycle; zooplankton communities are changing
-   Warming increases herring larval encounters with stressful or lethal surface temperature
:::

:::{
\vspace{0.5cm}
\section{NEW ENGLAND RISK POLICY SUMMARY (PLACEHOLDER)}
:::

::: {}
\vspace{0.5cm}

\raggedright

<!-- According to the [New England `r params$terminal_year` risk policy](`r params$risk_url`), Black Sea Bass scored high and/or moderately high risk in the following elements: -->

-   What type of information is useful to summarize here?
    -   Risks to meeting management objectives
    -   Compile existing risk indicators relevant to the stock
    -   Sources of management and model uncertainty
-   Some capacity to develop our own risk indicators and/or provide additional context based on other ecosystem risk indicators
-   Is there other information related to stock-level risk that should go here?
    
:::

:::{
\hfill
:::

::: {}
![](`r params$img2`)
<!-- ![](`r params$img3`) -->
:::

::::

\newpage

\newgeometry{top=0.75in, left=0.25in, right=0.25in, bottom=0.75in}

```{r}
data <- NEesp2::format_tbl_data(file = params$tbl_file,
                            term_year = params$terminal_year,
                            dir = params$img_dir)
```

```{r}
tbl <- NEesp2::rpt_card_table(data,
                       widths = params$widths)

tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "body")
tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "header")

tbl
```


::: {}

\vspace{-0.4cm}
\footnotesize \* The y-axis units are included in the "Indicator" column of the table. In all figures, the dashed line represents the time series mean, and the solid green lines indicate ± 1 standard deviation. Commercial data were derived from the commercial dealer database hosted at the Greater Atlantic Regional Office. All dollar values have been adjusted to 2024 real dollars. \newline


\centering\normalsize
We welcome your observations! Please contact northeast.ecosystem.highlights@noaa.gov with any on-the-water
insights or changes observed in the Atlantic herring fishery and [nefsc.esp.leads@noaa.gov](nefsc.esp.leads@noaa.gov) with questions or comments on the information presented in this report.
:::

\newpage
\newgeometry{top=1in, left=1in, right=1in, bottom=1in}

::: {}
\section{References}
1.  A. Molina (2025) Atlantic herring Ecosystem and Socioeconomic Profile
Trade-offs in the use of a seabird diet-derived recruitment index for the assessment of Atlantic herring in the Northeast U.S.

2.  D. Micah (2025) Development of a haddock predation index as a recruitment covariate in the Atlantic herring stock assessment model
3.  S. Gaichas (2024).Working Paper: Developing zooplankton indices in VAST food availability 

4.  A. Molina (2025) Temperature Threshold Indicators

5.  S. Gaichas & J. Deroba (2024) Working Paper: Recruitment covariate testing in WHAM

6.  J. Cournane (2024) ATLANTIC HERRING RESEARCH TRACK WORKING GROUP STAKEHOLDER ENGAGEMENT SESSION

7.  Center, C. P. (2024). North Atlantic Oscillation, Climate Prediction Center, National Centers for Environmental Predictions, NOAA [data set]. https://www.cpc.ncep.noaa.gov/products/precip/CWlink/pna/nao.shtml

8.  https://apps-garfo.fisheries.noaa.gov/quota-monitoring/atlantic-herring/archive-herring//2024-12-14_qm_herring.html

:::